Emergence of the Southeast Asian islands as a driver for Neogene cooling

### 2020-04-08 Manuscript/Manuscript.tex
Updates from Overleaf

\section*{Methods}

\subsection*{GEOCLIM model}

@@

\subsection*{Neogene cooling}

@@

\subsection*{Marine isotopes}

Himalayan uplift \cite{Raymo1992a, Galy2007a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling. Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}. This increase in \SrSr since ca. 35~Ma is followed by a decrease in slope ca. 15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision. Enhanced organic matter burial in the Bay of Bengal due to high sedimentation rates could also have contributed to Neogene cooling \cite{Galy2007a}.

@@

\subsection*{GEOCLIM silicate weathering component}

@@

>>>>

\section*{Model and Boundary Conditions}

\subsection*{GEOCLIM}

@@

\subsection*{SEAI Emergence}

@@

\subsection*{Volcanic Degassing}

\begin{figure}[h]
    \centering
    \includegraphics[width=1\linewidth]{Manuscript/Figures/weatherability_curves.pdf}
    \caption{Weatherability curves for the ``paleo-SEAI'' scenarios shown in Figure \ref{fig:scenario_pCO2}. Details on how these curves were generated are described in \MM. Each of the 4 curves represent a different tectonic boundary condition (i.e. the reconstructed paleoshorelines of the SEAI; Fig. \ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo (and vice versa). Point B represents the present day, in which \pCOtwo is 286~ppm. The arrow from Point A$_{1}$ to B represents the ``weatherability only'' scenario, in which global weatherability changes as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y. In this scenario, the \pCOtwo decreases from the value dictated by the 15~Ma ``paleo-SEAI'' weatherability curve ($\sim$570~ppm). The arrow from Point A$_{2}$ to B instead represents the ``degassing only'' scenario, in which global weatherability remains the same as the present day, but the same change in \pCOtwo as the ``weatherability only'' scenario is achieved by decreasing the volcanic degassing flux. In this scenario, the volcanic degassing flux decreases from the value dictated by the present day weatherability curve ($\sim$13\% greater than the present day). The arrow from Point A$_{3}$ to B represents the ``weatherability and degassing'' scenario, in which global weatherability changes as the SEAI emerge and the volcanic degassing flux decreases from a value 10\% greater than the present day.}
    \label{fig:weatherability_curves}
\end{figure}

A decrease in volcanic degassing \cite{Berner1983a} has also been proposed as a driver for Neogene cooling. Proxy-based estimates of the evolution of volcanic degassing fluxes throughout the Neogene are inconsistent with each other, such that not even the sign of the change in volcanic degassing over the past $\sim$15~m.y. is without ambiguity \cite{Godderis2017c}. For example, it has been estimated that the volcanic degassing flux was $\sim$25\% lower \cite{Cogne2006a} and $\sim$10\% higher \cite{Van-Der-Meer2014a} at 15~Ma relative to the present day.

Nevertheless, our model framework provides us with the opportunity to estimate the magnitude of volcanic degassing flux decrease necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.

\subsection*{Marine Isotopes}

Himalayan uplift \cite{Raymo1992a, Galy2007a} has also been proposed as drivers for Neogene cooling. Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}. This increase in \SrSr since ca. 35~Ma is followed by a decrease in slope ca. 15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision. Enhanced organic matter burial in the Bay of Bengal due to high sedimentation rates could also have contributed to Neogene cooling \cite{Galy2007a}.

@@

\subsection*{GEOCLIM Silicate Weathering Component}

@@

\subsection*{Weatherability Curves}

To create the 15~Ma ``paleo-SEAI'' curve shown in Figure \ref{fig:weatherability_curves}, we use the reconstructed paleoshorelines of the SEAI at 15~Ma (Fig. \ref{fig:shoreline_growth}A). We then select the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI), and fix \pCOtwo at the 3 \pCOtwo levels at which the GFDL CM2.0 climate model experiments were computed (\SI). We then run GEOCLIM at each of these \pCOtwo levels until steady state is achieved (i.e. until the volcanic degassing flux is equal to the silicate weathering flux). We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' scenarios and the present day to generate the 3 other weatherability curves presented in Figure \ref{fig:weatherability_curves}.
